### 2012-05-29 R_codes/turk1.Rnw
lmer with offset

<<read-data, echo=FALSE, resuls=hide>>=

suppressMessages(library(ggplot2))
setwd("/Users/heike/papers/paper-inference-jasa/R_codes/")
raw.dat <- read.csv("../data/raw_data_turk1.csv")
@

# ========== policy for excluding bad data =========
# Everyone is given at least one easy lineup. If a person can't correctly
# evaluate 50\% of those extremely easy lineups, all his responses are discarded.
# A lineup with p_value < 0.0002 is considered easy

# where are our reference charts?
# we should not draw inference on charts that we use for study inclusion

<<select, dependson=read-data, echo=FALSE, resuls=hide>>=
raw.dat$response <- raw.dat$response_no == raw.dat$plot_location

raw.dat$effect <- with(raw.dat, beta/sigma*sqrt(sample_size))

raw.dat$gender <- factor(raw.dat$gender, exclude=NULL)
levels(raw.dat$gender) <- c("male", "female", "unknown")
raw.dat$mage <- is.na(raw.dat$age)
raw.dat$meduc <- is.na(raw.dat$academic_study)

raw.dat$academic_study <- addNA(factor(raw.dat$academic_study))
raw.dat$age <- addNA(factor(raw.dat$age))

raw.dat$mgender <- raw.dat$gender=="unknown"

#dp <- ddply(subset(raw.dat, p_value < 0.0002),.(id),summarise,
#   percent_correct = mean(response)*100
#  )

#included_id <- dp$id[ dp$percent_correct > 0]
#dat <- subset(raw.dat, id %in% included_id)
dat <- raw.dat

suppressMessages(require(plyr))
turker <- ddply(dat,.(id), summarise,
	n = length(id),
	correct = sum(response),
  gender=gender[1]
)

## if we include all correct percentages above 0.2 we get rid of any differences due to gender, age or schooling
included <- unique(turker$id[with(turker, correct/n >= 0.1)])

## if we exclude everybody with just a single lineup, we have the difference introduced by missing values show up massively
#included <- unique(turker$id[turker$n > 1])

## if we exclude everybody with just a single lineup, we have the difference introduced by missing values show up massively
#included <- unique(turker$id[turker$gender != "unknown"])

dat <- subset(raw.dat, id %in% included)

lineups <- ddply(dat,.(pic_id), summarise,
	n = length(id),
	correct = sum(response),
  p_value = mean(p_value),
  beta=beta[1],
  sigma=sigma[1],
  sample_size=sample_size[1],
  effect = effect[1]               
)
@

<<fig-basic, dependson=select, echo=FALSE>>=
qplot(effect, p_value, data=lineups, colour=correct/n)
qplot(effect, correct/n, data=lineups, colour=p_value) + geom_smooth(method="loess")
@

<<model0, dependson=select>>=
suppressMessages(library(lme4))
dat$setup <- with(dat, sprintf("%d:%d:%d", beta, sigma, sample_size))

m0 <- lmer(response~ 1 + (1|id) + (1|setup), data=dat, family=binomial())
q01 <- qplot(ranef(m0)[[1]][,1], binwidth=0.2)
q02 <- qplot(ranef(m0)[[2]][,1], binwidth=0.2)
summary(m0)
summary(ranef(m0)[[1]][,1])
summary(ranef(m0)[[2]][,1])

dat$offset <- log(0.05/.95)
base <- lmer(response~ offset(offset) - 1 + effect + (effect|id) , data=dat, family=binomial())
anova(m0, base)
q11 <- qplot(ranef(base)[[1]][,1], binwidth=0.2)
q12 <- qplot(ranef(base)[[2]][,1], binwidth=0.2)
summary(base)

base.1 <- lmer(response~ effect + (effect|id) , data=dat, family=binomial())

m1 <- lmer(response~ effect + (1|id), data=dat, family=binomial())
anova(base, m1)

m1a <- lmer(response~ effect + gender + (1|id), data=dat, family=binomial())
m1b <- lmer(response~ effect + academic_study + (1|id), data=dat, family=binomial())
m1c <- lmer(response~ effect + age + (1|id), data=dat, family=binomial())

m2a <- lmer(response~ effect + mgender + (1|id), data=dat, family=binomial())

anova(m0, base,  m2a, m1a)
anova(base, m1,  m2a, m1b)
anova(base, m2a, m1c)

dat$alpha <- 0.05
m3c <- lmer(response~ offset(alpha)+effect -1 + age + (effect|id), data=dat, family=binomial())
anova(base, m2a, m1c, m3c)

# there's something in model m1c that's not explained by missingness alone - but it might be an artefact - there's nothing immediately visible in the estimates
@

<<plots, dependson=model0, echo=FALSE, results='hide', out.width='\\text.width'>>=
qplot(ranef(m3c)[[1]][,1], binwidth=0.1) + xlab("Individuals' Skill")

qqnorm(ranef(m1c)[[1]][,1])
abline(a=0,b=1)

qplot(effect, fitted(base), data=dat, colour=difficulty,  group=factor(sample_size)) + geom_smooth()

qplot(effect, fitted(m1c), data=dat, colour=factor(sample_size))
qplot(group=effect, x=resid(m1c), geom="histogram", data=dat, facets=~effect, binwidth=0.2)
@

levels(dat$age) <- c("below 18", "18 - 25", "26 - 30", "31 - 35", "36 - 40", "41 - 45", "46- 50", "51 - 55", "56 - 60", "above 60")

# academic_study 1 = High school or less
                 2 = Some under graduate courses 
                 3 = Under graduate degree
                 4 = Some graduate courses
                 5 = Graduate degree

# conf_level 1 = most, 5 = least

# choice_reason 1 = Big vertical difference 
                2 = Medians are different
                3 = Outliers
                4 = others

# unit(time_taken) = second

<<predict, dependson=model0>>=
predict.lme <- function(model, newdata, skills=0) {
  design <- ldply(newdata$age, function(x) levels(model@frame$age) %in% x)
  xnew <- cbind(effect=newdata$effect, design)
  preds <- drop(as.matrix(xnew) %*% fixef(model)) + skills
  exp(preds)/(exp(preds)+1)
}

newdata <- data.frame(expand.grid(effect=seq(0,17, by=0.1), age=c(0:10, NA)))
newdata$m1c <- predict.lme(m3c, newdata=newdata)

newdata_all <- data.frame(expand.grid(effect=seq(0,17, by=0.1), age=c(0:10, NA), skills=ranef(m1c)[[1]][,1]))
newdata_all$m1c <- predict.lme(m1c, newdata=newdata_all, skills=ranef(m1c)[[1]][,1])

qplot(effect, m1c, data=newdata, colour=factor(age))
qplot(effect, m1c, data=newdata_all, colour=factor(age))
@

# ------function to calculate UMP power -----------------------

calculate_ump_power <- function (beta, n, sigma){
	alpha <- 0.05/2
	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs for derivation of power
	mu <- beta/se_beta
	t_n <- qt(p=1-alpha,df=n-3)
	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
	return(res)
}

calculate_ump_power(3,100,5)
calculate_ump_power(3,300,5)

calculate_ump_power(3/5,300,1)

# ---------------------- fitting loess smoother  ------------

fit <- loess(as.numeric(response) ~ beta,data=subset(dat,sigma==5 & sample_size==100),span=1.2)
betas <- seq(0,9,by=.1)
with(subset(dat,sigma==5 & sample_size==100),qplot(betas,predict(fit,data.frame(beta=betas))))

fit <- loess(as.numeric(response) ~ beta,data=subset(dat,sigma==5 & sample_size==300),span=.75)
betas <- seq(0,5,by=.1)
with(subset(dat,sigma==5 & sample_size==300),qplot(betas,predict(fit,data.frame(beta=betas))))

get_smooth_power <- function(dat.n,test="Empirical"){
 betas <- seq(0.01,16, by=.2)
 dat_smooth <- NULL
 for (s in c(5,12)){
	for (n in c(100,300)){
		dats <- subset(dat.n,sigma==s & sample_size==n)
		dats$y <- as.numeric(dats$response)
		fit <- loess(y ~ beta,data=dats,span=1+(n==100)*.2)
		dat_smooth <- rbind(dat_smooth,cbind(betas,predict(fit,betas),n,s))
		}
	}
 colnames(dat_smooth) <- c("beta","pow","sample_size","sigma")
 dat_empirical <- data.frame(rbind(dat_smooth,cbind(-dat_smooth[,1],dat_smooth[,-1])))
 dat_empirical$test <- test
 return(dat_empirical)
}

get_ump_power <- function(dat,test="UMP"){
 beta <- seq(0.01,16, by=.2)
 dat_pow <- NULL
 for(n in c(100,300)){
  for(sg in c(5,12)){
     pow <- NULL
     for (i in beta)pow <- c(pow,calculate_ump_power(beta=i,n=n,sigma=sg))
     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))
     dat_ump$sample_size <- n
     dat_ump$sigma <- sg
     #head(dat_ump)
     dat_pow <- rbind(dat_pow,dat_ump)
  }
 }
 dat_pow$test <- test	
 return(dat_pow)
}

dat_emp_pow <- get_smooth_power(dat)
dat_ump_pow <- get_ump_power(dat)

p <- ggplot(dat_emp_pow,aes(beta,pow)) + geom_line(aes(colour=test)) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power")
p

# ---------- bootstrap band for empirical power -----------------------

dat_boot_pow <- NULL
for (i in 1:1000){
dat.b <- ddply(dat,.(beta,sample_size,sigma), summarize,
           response = sample(response,replace=T)
           )
dat_boot_pow <- rbind(dat_boot_pow,get_smooth_power(dat.b,test=paste("smooth",i,sep="")))
}

#write.csv(dat_boot_pow,file="../data/dat_bootstrap_power.txt",row.names=F)
#dat_boot_pow <- read.csv("../data/dat_bootstrap_power.txt")

p <- ggplot(dat_boot_pow, aes(beta,pow,colour=grey)) + geom_line(aes(group=test),alpha=I(0.1))
p + facet_grid(sample_size~sigma)

p <- ggplot() +
     geom_line(aes(beta,pow, group =test), data=dat_boot_pow,alpha=I(0.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_boot_limit <-  ddply(dat_boot_pow,.(beta,sample_size,sigma), summarize,
           limit1 = quantile(pow,.025, na.rm=T),
           limit2 = quantile(pow,.975, na.rm=T)
           )

dat_boot_pow_limit <- melt(dat_boot_limit, id=c("beta","sample_size","sigma"))
colnames(dat_boot_pow_limit) <- c("beta","sample_size","sigma","test","pow")

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,as.numeric(response)), data=dat) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_obs_val <- ddply(dat,.(beta,sample_size,sigma,response), summarize,
           responses = length(response)
           )
dat_obs_val <- rbind(dat_obs_val,cbind(beta= -dat_obs_val[,1],dat_obs_val[,-1]))

p <- ggplot() +
     geom_line(aes(beta,pow, group =test), data=dat_boot_pow,alpha=I(0.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_obs_pow <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,
           pow = sum(response=="TRUE")/sum(response==response)
           )
dat_obs_pow <- rbind(dat_obs_pow,cbind(beta= - dat_obs_pow[,1],dat_obs_pow[,-1]))

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,pow), data=dat_obs_pow, colour=alpha("red",.3)) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_boot_ribbon <-  ddply(dat_boot_pow,.(beta,sample_size,sigma), summarize,
           limit1 = quantile(pow,.025, na.rm=T),
           limit2 = quantile(pow,.975, na.rm=T)
           )
dat_boot_ribbon <- rbind(dat_boot_ribbon,cbind(beta= -dat_boot_ribbon[,1],dat_boot_ribbon[,-1]))

p <- ggplot() +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val, colour=alpha("red",.3)) +
     geom_ribbon(aes(x=beta,ymin=limit1,ymax=limit2), data=dat_boot_ribbon, fill=alpha("red",.2)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

p <- ggplot() +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val, colour=alpha("black",.3)) +
     geom_ribbon(aes(x=beta,ymin=limit1,ymax=limit2), data=dat_boot_ribbon, fill=alpha("grey",.9)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

ggsave(p, filename="../images/power_loess_exp1.pdf", height=5.5, width=8.5)

# --------------- p_value vs %correct ----------------------------------

pdat <- ddply(dat, .(pic_name), summarize,
   p_value = mean(p_value),
	attempted = sum(response==response),
	corrected = sum(response=="TRUE"),
	percent_correct = sum(response=="TRUE")*100/sum(response==response)
	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat) +
     geom_point(aes(p_value,percent_correct)) + 
     xlab(expression(paste("p-value(",p[B],")"))) +
     ylab("Percentage of correct responses") 
p

ggsave("../images/p_val_percent_correct.pdf")

pdat$lbl <-""
indx <- with(pdat,(percent_correct>30 & p_value>.05) | (percent_correct>5 & p_value>.5))
pdat$lbl[indx] <- as.character(pdat$pic_name[indx])
pdat$angle <- 0
pdat$angle[pdat$percent_correct>5 & pdat$p_value>.5] <- 90

p <- ggplot(pdat,aes(p_value,percent_correct,label=lbl)) +
     geom_point() + geom_text(hjust=-.05,vjust=0,aes(angle=angle)) +
     xlab(expression(paste("p-value(",p[B],")"))) +
     ylab("Percentage of correct responses") +
     geom_vline(xintercept=0.05) + geom_hline(yintercept=5) +geom_smooth()
p

subset(dat,pic_name=="plot_turk1_100_1_5_1.png")
subset(dat,pic_name=="plot_turk1_100_0_5_1.png")
subset(dat,pic_name=="plot_turk1_300_1_12_1.png")

subset(dat,pic_name=="plot_turk1_100_1_5_3.png")
subset(dat,pic_name=="plot_turk1_100_1_12_3.png")

# ------------- observed and theorectical power curves ------------------------------

calculate_theoretical_power <- function (beta, n, sigma){

	# n <- 100

	# sigmasq <- 12^2

	alpha <- 0.05

	sigmasq <- sigma^2

	beta_not <- 0

	se_beta <- sqrt(sigmasq/(n*(.5^2)))    # refer to docs derivation of power

	mu <- beta/se_beta

	alpha <- alpha/2

	t_n <- qt(p=1-alpha,df=n-3)

	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1

	return(res)

}

exact_ci <- function(yy, nn, alpha=0.05) {
	pL <- rep(0,length(yy))
	pU <- rep(1,length(yy))

	y <- yy[yy!=0]
	n <- nn[yy!=0]
	pL[yy!=0] <- 1/(1+ (n-y+1)/(y*qf(alpha/2, 2*y, 2*(n-y+1))))

	y <- yy[yy!=nn]
	n <- nn[yy!=nn]
	pU[yy!=nn] <- 1/(1+ (n-y)/((y+1)*qf(1-alpha/2, 2*(y+1), 2*(n-y))))
	pw <- yy/nn
	return(data.frame(pw,pL, pU))
}
exact_ci(yy=c(3,4),nn=c(5,7))

dat_obs <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,
           empirical_power = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,1]
           )
dat_obs1 <- dat_obs

power_obs_ump <- cbind(dat_obs[,1:4],ump=sapply(dat_obs[,1],calculate_theoretical_power ,n=100,sigma=5))
#setwd("U:/phd research/simulation_study/turk")
#write.table(power_obs_ump,file="power_obs_ump_exp1.txt",row.names=F)

dat_obs1$beta <- -dat_obs1$beta
dat_obs <- rbind(dat_obs1,dat_obs)
dat_obs.m <- melt(dat_obs, id=c("beta","sample_size","sigma", "replica"))
colnames(dat_obs.m) <- c("beta","sample_size","sigma", "replica","test","pow")
head(dat_obs.m)

p <- qplot(beta,pow, geom=c("point","line"), data=dat_obs.m, colour=factor(replica)) 
p + facet_grid(sample_size~sigma)

p <- qplot(beta,pow, geom=c("smooth","point"), data=dat_obs.m) 
p + facet_grid(sample_size~sigma)

p <- ggplot(dat_obs.m, aes(beta,pow)) + geom_point()+geom_smooth()
p + facet_grid(sample_size~sigma)

=c("smooth","point"), data=dat_obs.m) 
p + facet_grid(sample_size~sigma)

beta <- seq(0.01,16, by=.2)
dat_pow <- NULL
for(n in c(100,300)){
  for(sg in c(5,12)){
     pow <- NULL
     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))
     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))
     dat_ump$sample_size <- n
     dat_ump$sigma <- sg
     #head(dat_ump)
     dat_pow <- rbind(dat_pow,dat_ump)
}}
dat_pow$test <- "UMP"
head(dat_pow)

plot_dat <- rbind(dat_obs.m,dat_pow[,c(1,3,4,5,2)])
head(plot_dat)
tail(plot_dat)

p <- qplot(beta,pow, geom="line", data=plot_dat, colour=test, linetype=test) 
p <- p + facet_grid(sample_size~sigma)
p <- p + scale_colour_manual(values=c("red","red","red","cyan4"))
p <- p + scale_linetype_manual(values=c(1,2,2,1))
p + xlab(expression(beta)) + ylab("Power")

p <- ggplot(dat_obs.m, aes(beta,pow, colour=test)) + geom_point(alpha=0.5)+geom_smooth()
p <- p + geom_line(aes(beta,pow, colour=test), data=dat_pow)
p <- p + facet_grid(sample_size~sigma)
p <- p + xlab(expression(beta)) + ylab("Power")

ggsave(p, filename="../images/emperical_ump_power_exp1.pdf", height=5.5, width=8.5)

# ----------------- glm power curve ----------------------------------

my_glm_predict <- function(dat.new){

lb <- function(b){

   y <- as.numeric(dat$response)-1

   b0 <- log(0.05/(1-0.05))

   x <- with(dat,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma, x4=beta*sample_size*sigma))

   xb <- b0 + x%*%b

   p <- exp(xb)/(1+exp(xb))

   return(-sum(y*log(p)+(1-y)*log(1-p)))

}

#lb(c(.2,.002,.02,.0002))

bh <- optim(par=c(.2,.002,.02,.0002),lb)$par

xnew <- with(dat.new,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma, x4=beta*sample_size*sigma))

xbh <- b0 + xnew%*%bh

ph <- exp(xbh)/(1+exp(xbh))

return(ph)

}

beta <- rep(seq(.1,16, by=.25),4)

sample.size <- c(rep(100,length(beta)/2),rep(300,length(beta)/2))

sigma <- rep(c(12,5,12,5),each=length(beta)/4)

dat.new <- data.frame(beta, sample_size=sample.size,sigma)

power <- my_glm_predict(dat.new)

vdat <- data.frame(beta=c(beta,-beta),power=c(power,power),sample.size=rep(sample.size,2),sigma=c(sigma,sigma))

vdat$Test="Visual"

qplot(beta,power, geom="line", data=vdat)+facet_grid(sample.size~sigma)

qplot(beta,power, geom="line", data=subset(vdat,(sigma==12)&(sample.size==100)))

udat <- NULL

for(n in c(100,300)){

  for(sg in c(5,12)){

     pow <- NULL

     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))

     dat_ump <- data.frame(beta=c(-beta,beta),power=c(pow,pow))

     dat_ump$sample.size <- n

     dat_ump$sigma <- sg

     #head(dat_ump)

     udat <- rbind(udat,dat_ump)

}}

udat$Test <- "UMP"

dat.pow <- rbind(vdat,udat)

qplot(beta,power, geom="line", data=dat.pow, colour=Test)+facet_grid(sample.size~sigma)

# ------------------ power curve using glm function -----------------------

fit.power <- glm(response ~ sample_size+beta+sigma

             , family=binomial,data=dat)

res <- summary(fit.power)

str(res)

res$coef

beta <- seq(.1,16, by=.25)

sample.size <- rep(100,length(beta))

sigma <- rep(12,length(beta))

newdat <- data.frame(sample_size=sample.size,sigma,beta)

power <- predict(fit.power, newdata = newdat, type="response", se.fit = TRUE)

pr <- predict(fit.power, newdata = newdat, se.fit = TRUE)

family <- family(fit.power)

lower <- family$linkinv(pr$fit - qnorm(0.95) * pr$se.fit)

upper <- family$linkinv(pr$fit + qnorm(0.95) * pr$se.fit)

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , empirical=rep(power$fit,2)

                    , lower_limit=rep(lower,2),upper_limit=rep(upper,2)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

colnames(pow.dat.m) <- c("betas","Test","power")

p <- qplot(betas, power, geom="line", data=pow.dat.m, color=Test, linetype=Test)

p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(1,2,2,1))

# ---------------- estimating power curves for each replicate of linup ----------------

my_glm_predict.rep <- function(dat.new){

  b0 <- log(0.05/(1-0.05))

  lb <- function(b){

     y <- as.numeric(dat$response)-1

     x <- with(dat,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma,

             x4=replica*beta,x5=beta*sample_size*sigma*replica))

     xb <- b0 + x%*%b

     p <- exp(xb)/(1+exp(xb))

     return(-sum(y*log(p)+(1-y)*log(1-p)))

  }

  #lb(c(.2,.007,.002,-.0002,-.002))

  bh <- optim(par=c(.2,.007,.002,-.0002,-.002),lb)$par

  # 0.2162250624  0.0056327718 -0.0174778836  0.0773441682 -0.0001383814

  xnew <- with(dat.new,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma,

             x4=replica*beta,x5=beta*sample_size*sigma*replica))

  xbh <- b0 + xnew%*%bh

  ph <- exp(xbh)/(1+exp(xbh))

  return(ph)

}

# ----- investigating ------

matrix(my_glm_predict.rep(subset(dat.new, sample_size==100 & sigma==5)),ncol=3)

# ----- end investigation -------

beta <- seq(.1,16, by=.25)

sample.size <- c(100,300)

sigma <- c(12,5)

replica <- 1:3

dat.new <- expand.grid(beta,sample.size,sigma,replica)

colnames(dat.new) <- c("beta","sample_size","sigma","replica")

dat.new$power <- my_glm_predict.rep(dat.new)

vdat <- with(dat.new,data.frame(beta=c(beta,-beta),power=c(power,power),

        sample.size=rep(sample_size,2),sigma=c(sigma,sigma), replica=c(replica,replica)))

vdat$Test=paste("Replicate",vdat$replica,sep="")

qplot(beta,power, geom="line", data=vdat, colour=Test)+facet_grid(sample.size~sigma)

qplot(beta,power, geom="line",colour=Test,data=subset(vdat,(sigma==5)&(sample.size==100)))

dat.X <- expand.grid(beta,sample.size,sigma)

colnames(dat.X) <- c("beta","sample.size","sigma")

upow <- apply(dat.X,1,function(x){return(calculate_theoretical_power(beta=x[1],n=x[2],sigma=x[3]))})

udat <- data.frame(beta=c(dat.X$beta, -dat.X$beta),

                  power=c(upow,upow), rbind(dat.X[,-1],dat.X[,-1]),

                  replica=0, Test="UMP")

dat.pow <- rbind(vdat,udat)

p <- qplot(beta,power, geom="line", data=dat.pow, colour=Test, linetype=Test)

p <- p + facet_grid(sample.size~sigma)

p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

dat_obs <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,

	power=sum(response=="TRUE")/length(response)

   )

dat_obs$sample.size=dat_obs$sample_size

dat_obs$Test <- paste("Replicate",dat_obs$replica, sep="")

ggplot()+

geom_line(data=dat.pow,aes(x=beta,y=power,colour=Test, linetype=Test)) +

geom_point(data=dat_obs,aes(x=beta,y=power,colour=Test)) +

facet_grid(sample.size~sigma) +

xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

dat.pow.diff <- vdat

dat.pow.diff$power <- rep(udat$power,3)- vdat$power

p <- qplot(beta,power, geom="line", data=subset(dat.pow.diff,beta>=0), colour=Test, linetype=Test)

p <- p + facet_grid(sample.size~sigma)

p + xlab(expression(beta)) + ylab("Difference from UMP test Power") + scale_linetype_manual(values =c(2,3,4,1))

# ---------------- estimating power curves for each replicate of linup using glm ----------------

dat$bf <- dat$beta*dat$replica

fit.power.rep <- glm(response ~ sample_size+sigma+bf

             , family=binomial,data=dat)

res <- summary(fit.power.rep)

str(res)

res$coef

dat.pw <- NULL

for ( nn in c(100,300)){

for (sg in c(5,12)){

beta <- seq(.1,16, by=.25)

nb <- length(beta)

sample.size <- rep(nn,nb)

sigma <- rep(sg,nb)

newdat <- data.frame(sample_size=sample.size,sigma)

newdat <- rbind(newdat,newdat,newdat)

newdat$bf  <- rep(beta,3)* rep(1:3, each=nb)

power.rep <- predict(fit.power.rep, newdata = newdat, type="response", se.fit = TRUE)

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=nn,sigma=sg))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , replicate1=rep(power.rep$fit[1:nb],2)

                    , replicate2=rep(power.rep$fit[(nb+1):(nb*2)],2)

                    , replicate3=rep(power.rep$fit[(2*nb+1):(nb*3)],2)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

colnames(pow.dat.m) <- c("betas","Test","power")

#p <- qplot(betas, power, geom="line", data=pow.dat.m, color=Test, linetype=Test)

#p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

replicate1 <- ump_pow-power.rep$fit[1:nb]

replicate2 <- ump_pow-power.rep$fit[(nb+1):(nb*2)]

replicate3 <- ump_pow-power.rep$fit[(2*nb+1):(nb*3)]

overall <- ump_pow-power$fit

pow.diff <- data.frame(beta,replicate1,replicate2,replicate3,overall)

pow.diff.m <- melt(pow.diff,id="beta")

colnames(pow.diff.m) <- c("betas","Replicate","power")

#p <- qplot(betas, power, geom="line", data=pow.diff.m, color=Replicate, linetype=Replicate)

#p + xlab(expression(beta)) + ylab("Difference from UMP Power") + scale_linetype_manual(values =c(4,3,2,1))

dat.pw <- rbind(dat.pw,cbind(pow.diff.m,sample_size=nn,sigma=sg))

}}

p <- qplot(betas, power, geom="line", data=dat.pw, color=Replicate, linetype=Replicate)

p <- p + xlab(expression(beta)) + ylab("Difference from UMP Power")

p + scale_linetype_manual(values =c(4,3,2,1)) + facet_grid(sample_size~sigma)

# ----------------- estimating subjectwise estimated power curve -------------------------

library(lme4)

fit.mixed <- lmer(response ~ factor(sample_size)+beta+factor(sigma)

              + (1|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#res@frame

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

X <- cbind(rep(1,length(beta)),sample.size,beta,sigma.val)

dim(X)

delta <- as.numeric(res@REmat[4])

power=NULL

M <- 30

for (i in 1:M){

tau <- rnorm(mean=0,sd=delta, n=1)

xb <- X %*% B + tau

power <- cbind(power,exp(xb)/(1+exp(xb)))

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=I(alpha("black", 0.1)), data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="hotpink", size=1)

# ----------------- subjectwise power curve with random slope ---------------

library(lme4)

fit.mixed <- lmer(response ~ factor(sample_size)+beta+factor(sigma)

              + (1+beta|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#res@frame

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

ones <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

X <- cbind(ones,sample.size,beta,sigma.val)

dim(X)

Z <- cbind(ones,beta)

cov <- as.numeric(res@REmat[2,5])*as.numeric(res@REmat[1,4])*as.numeric(res@REmat[2,4])

vv <- matrix(as.numeric(c(res@REmat[1,3],rep(cov,2),res@REmat[2,3])),ncol=2, byrow=F)

power=NULL

M <- 20

#library(MASS)

for (i in 1:M){

tau <- mvrnorm(n = 1, mu=c(0,0),Sigma=vv)

xb <- X %*% B + Z %*%tau

power <- cbind(power,exp(xb)/(1+exp(xb)))

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=I(alpha("grey", 1)), data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="hotpink", size=1)

# ----------------- estimating subjectwise power curve by gender etc.------------------

library(lme4)

fit.mixed <- lmer(response ~ gender+factor(sample_size)+beta+factor(sigma)

              + (1|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#head(res@frame)

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

delta <- as.numeric(res@REmat[4])

power=NULL

gnd <- NULL

tau <- res@ranef    # estimates for existing subject

M <- length(tau)

for (i in 1:M){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,length(beta))

  X <- cbind(rep(1,length(beta)),gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + tau[i]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1:M, each=length(beta)*2)],rep(1,length(beta)*2)), labels=c("Male","Female"))

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="maroon4", size=1)

# ----------------- estimating subjectwise power curve by gender and random slope------------------

library(lme4)

fit.mixed <- glmer(response ~ gender+factor(sample_size)+beta+factor(sigma)

              + (1 + beta|id)

              , family=binomial

              , data=dat)

res <- summary(fit.mixed)

B <- fixef(fit.mixed)

u <- as.matrix(ranef(fit.mixed)$id)

head(u)

beta <- seq(0.01,16, by=.2)

nb <- length(beta)

sample.size <- rep(1,nb)

sigma.val <- rep(0,nb)

ones <- rep(1,nb)

#Z <- cbind(ones,beta,sample.size,sigma.val)

#Z <- cbind(ones,beta,sigma.val)

Z <- cbind(ones,beta)

power=NULL

gnd <- NULL

M <- length(u[,1])

s <- sample(1:M,size=20)

#s <- c(213,8)

for (i in s){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,nb)

  X <- cbind(ones,gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + Z %*% u[i,]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- s

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

#head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1:length(s), each=nb*2)],rep(1,nb*2)), labels=c("Male","Female"))

#head(pow.dat.m)

#tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="maroon4", size=1)

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="darkviolet", size=1)

# ----------------- individual observed and estimated power curve by gender and random slope----------

power=NULL

gnd <- NULL

M <- length(u[,1])

s <- sample(1:M,size=1)

for (i in s){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,nb)

  X <- cbind(ones,gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + Z %*% u[i,]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- s

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

#head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1, each=nb*2)],rep(1,nb*2)), labels=c("Male","Female"))

#head(pow.dat.m)

#tail(pow.dat.m)

dat_s=as.data.frame(subset(dat,id==s))

dat_s$variable=1

dat_pi <- merge(dat_s,ddply(dat_s,.(beta),summarize, obs_pi = mean(as.numeric(response)-1) ), by.beta=by.beta)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p <- p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="darkviolet", size=1)

p <- p + geom_point(aes(beta,as.numeric(response)-1), data=dat_s)

p + geom_point(aes(beta,obs_pi, colour="Observed"), data=dat_pi)

# ------------- observed and theorectical power curves ------------------------------

exact_ci <- function(yy, nn, alpha=0.05) {

	pL <- rep(0,length(yy))

	pU <- rep(1,length(yy))

	y <- yy[yy!=0]

	n <- nn[yy!=0]

	pL[yy!=0] <- 1/(1+ (n-y+1)/(y*qf(alpha/2, 2*y, 2*(n-y+1))))

	y <- yy[yy!=nn]

	n <- nn[yy!=nn]

	pU[yy!=nn] <- 1/(1+ (n-y)/((y+1)*qf(1-alpha/2, 2*(y+1), 2*(n-y))))

	pw <- yy/nn

	return(data.frame(pw,pL, pU))

}

exact_ci(yy=c(3,4),nn=c(5,7))

dat_obs <- ddply(dat,.(beta,sample_size,sigma), summarize,

               empirical = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,1],

               lower_limit = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,2],

               upper_limit = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,3]

              )

dat_obs1 <- dat_obs

power_obs_ump <- cbind(dat_obs[,1:4],ump=sapply(dat_obs[,1],calculate_theoretical_power ,n=100,sigma=5))

#setwd("U:/phd research/simulation_study/turk")

#write.table(power_obs_ump,file="power_obs_ump_exp1.txt",row.names=F)

dat_obs1$beta <- -dat_obs1$beta

dat_obs <- rbind(dat_obs1,dat_obs)

dat_obs.m <- melt(dat_obs, id=c("beta","sample_size","sigma"))

colnames(dat_obs.m) <- c("beta","sample_size","sigma","test","pow")

head(dat_obs.m)

qplot(beta,pow, geom="line", data=dat_obs.m, colour=test) + facet_grid(sample_size~sigma)

beta <- seq(0.01,16, by=.2)

dat_pow <- NULL

for(n in c(100,300)){

  for(sg in c(5,12)){

     pow <- NULL

     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))

     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))

     dat_ump$sample_size <- n

     dat_ump$sigma <- sg

     #head(dat_ump)

     dat_pow <- rbind(dat_pow,dat_ump)

}}

dat_pow$test <- "UMP"

head(dat_pow)

plot_dat <- rbind(dat_obs.m,dat_pow[,c(1,3,4,5,2)])

head(plot_dat)

tail(plot_dat)

p <- qplot(beta,pow, geom="line", data=plot_dat, colour=test, linetype=test)

p <- p + facet_grid(sample_size~sigma)

p <- p + scale_colour_manual(values=c("red","red","red","cyan4"))

p <- p + scale_linetype_manual(values=c(1,2,2,1))

p + xlab(expression(beta)) + ylab("Power")

# --------------- p_value vs time_taken ------------------------------

ggplot(dat,aes((p_value)^.2,(time_taken)^.1))+geom_point()+geom_smooth()

p <- ggplot(dat, aes(factor(p_value), time_taken))

p <- p + geom_boxplot() +ylim(0,450) + facet_grid(gender~.)

p

pdat1 <- ddply(dat, .(p_value, sample_size), summarize,

	avg_time_taken = median(time_taken),

	num_responded = sum(p_value==p_value),

      nn = length(p_value)

    )

p <- ggplot(subset(pdat1, p_value<= 0.2), aes(p_value,avg_time_taken))

p <- p+ geom_point(aes(size=nn))

p <- p + stat_smooth(se=F)

p + facet_grid(sample_size~.)

# --------------- Comparison of p_values from theory and actual ----------------------

pval <- function(x){

	N <- x[1]

	u <- x[2]

	k <- u:N

	binom <- choose(N,k)*((0.05)^k)*((0.95)^(N-k))

	return(sum(binom))

}

pval(c(13,3))

pdat <- ddply(dat, .(p_value, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

    )

pval_calculated <- apply(pdat[,3:4],1,pval)

qplot(pdat$p_value,pval_calculated, xlab="p_value of Mathematical Test", ylab="p_value of Visual Test")

ddply(dat, .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(subset(dat, pic_id==92), .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(subset(dat, substr(pic_name,1,12)=='plot_100_3_5'), .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(dat, .(pic), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# --------------- p_value vs %correct ----------------------------------

pdat <- ddply(dat, .(p_value, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat, aes(p_value,percent_correct))

p+ geom_point(size=2)

ggsave("../images/p_val_percent_correct.pdf")

p <- ggplot(pdat, aes(p_value,percent_correct))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(se=F)

p + facet_grid(sample_size~.)

p <- ggplot(subset(pdat, p_value <= 0.099), aes(p_value,percent_correct))

p <- p+ geom_point(size=3)

p <- p + stat_smooth( se=F)

p + facet_grid(sample_size~.)

# --------------- distribution of time taken ----------------------------------

p <- ggplot(dat, aes(time_taken))+geom_histogram(binwidth=2)+xlim(0,350)

p <- p + facet_grid(gender~.)

p

# --------------- p_value vs %correct by male female ----------------------------------

pdat <- ddply(dat, .(p_value, gender, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat, aes(p_value,percent_correct, colour=gender, shape=gender))

p+ geom_point(size=3)

p <- ggplot(pdat, aes(factor(p_value),percent_correct, colour=gender, shape=gender))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p + facet_grid(sample_size~.)

p <- ggplot(subset(pdat, p_value <= 0.099), aes(factor(p_value),percent_correct, colour=gender, shape=gender))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p + facet_grid(sample_size~.)

# ------------- number of attempts and % correct -----------

dat$attempt_no <- unlist(tapply(dat$start_time, dat$nick_name,

      function(x){return(as.integer(factor(x)))}), use.names=F)

adat <- ddply(dat, .(attempt_no, gender), summarize,

	attempted = sum(response_no==response_no),

	corrected = sum(response_no==obs_no),

      percent_correct = sum(response_no==obs_no)*100/sum(response_no==response_no)

    )

p <- ggplot(subset(adat, attempt_no <= 12), aes(attempt_no,percent_correct, color=factor(gender), shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)

# --------------- number of attempts vs time_taken ------------------------

adat1 <- ddply(dat, .(attempt_no, gender), summarize,

	avg_time_taken = mean(time_taken),

	avg_p_value = mean(p_value)

    )

p <- ggplot(subset(adat1, attempt_no <= 12), aes(attempt_no,avg_time_taken, color=gender, shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p <- ggplot(subset(adat1, attempt_no <= 12), aes(attempt_no,avg_p_value, color=gender, shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)