### 2012-08-08 cover-letter.tex
cover letter

\documentclass{letter}
\usepackage{graphics}
\usepackage{fullpage}
\topmargin=-0.2in
%\headheight=0in
\setlength\oddsidemargin{-0.2in}
\textheight=8.5in
\textwidth=7in
\def\apl#1#2{
\begin{letter}{#1}
\signature{Mahbubul Majumder\\
with Heike Hofmann and Dianne Cook\\Department of Statistics\\Iowa State University 
\\Ames, Iowa 50011-1210\\ Phone (515) 294 3440\\{\tt mahbub@iastate.edu}}
\opening{{Dear #2,}}

Please find enclosed a manuscript  ``Validation of Visual
Statistical Inference, Applied to Linear Models'' which we are
submitting for consideration for publication in the
Theory and Methods section of the Journal of American Statistical
Association.

This is the second foundational paper describing research which bridges
graphics, exploratory data analysis/data mining, and statistical
inference. The first paper by Buja, et al (2009) RSS Phil. Trans. A
set up the conceptual foundation. This paper clarifies and defines the
approach and validates it with a user study. There are two more papers
published on the topic, one that puts the ideas in the information
visualization context that won the best conference paper at InfoVis
2010, and another that explains how the method can be used to choose
the best graphic for a particular purpose, that is to be presented at
InfoVis 2012 in October.

The JASA instructions for authors say

\begin{quote}
``Authors are reminded that, although statistical theory has roots in mathematics, the phrase theory and methods should be interpreted broadly to include all techniques relevant to statistics and probability. This may include computational and graphical methods as well as more traditional mathematical methods.''
\end{quote}
We believe this paper is of interest to  the JASA readership. One concern we have is that it may be
considered to be long by JASA standards (28 pages, single spacing) and
we would not find it easy to shorten it.

Thank you for your consideration of our work.

\closing{Sincerely yours,
}
\end{letter}}

\begin{document}

{{\pdfimage width 4in {letterhead.jpg}}}
\apl{Professor Xuming He and Jun Liu\\Theory and Methods Editors\\Journal of American Statistical Association}{Professors He and Liu}

\end{document}